---
title: "MD1 Model Answer"
author: "Professor Vaisey"
output: 
  html_document:
    keep_md: true
---

# Question 1

First I will install some packages. Since I only want to do this one time (not every time I render this document) I am going to set the "chunk option" `eval=FALSE`. This tells R not to "evaluate" (i.e., execute) this block of code.

```{r,eval=FALSE}
install.packages("causact")
install.packages("dplyr")
install.packages("igraph")
```

# Question 2

Now that those are installed, I can load them.

```{r}
library(causact)
library(dplyr)
library(igraph)
```

I got the warning messages, as anticipated.

```{r}
# df <- as_data_frame(x = c(1, 2, 3))
```

I tried to run the code above but (as expected!) it gave me an error. It said that x "wasn't a graph object." I'm not sure what that means! For now, I commented out the code so I can render the document later.

```{r}
df <- dplyr::as_data_frame(x = c(1, 2, 3))
glimpse(df)
```

This worked but, as expected, it gave me a warning. It's telling me I should be using `as_tibble()` instead. I'm not sure exactly what that means, but at least I didn't get an error message!

As the homework said, putting the package name followed by two colons will make sure I get the function I want. So let me try this with another function.

```{r}
x <- c(5, 6, 2, 7, 9, 1)
dplyr::n_distinct(x)
```

OK this worked, but I'm not totally sure what it's doing at this point.

I think earlier R was using the `igraph` version of `as_data_frame()` instead of the `dplyr` version of `as_data_frame()` because it was "masked." I guess R uses the last-loaded version of a function so I have to be careful in which order I load packages! Or if I really want to be sure which version of a function I'm using, I'd better prefix it with the package name and two colons.

# Question 3

When I type `?n_distinct()` it brings up the help file for that function. Apparently it counts the number of unique values in a vector.

# Question 4

```{r}
glimpse(baseballData)
```

This dataset has 12,145 rows and 5 columns. The `Home` column is a "factor" type variable and `HomeScore` is an integer variable.

# Question 5

```{r}
baseballData[1,]
```

This is the first row of the `baseballData` dataset. It looks like it has information about a single game of baseball.

```{r}
baseballData[,2:3] |> head()
```

This function shows me two columns of the baseball data. The first is the acronym for the home team, the second is the acronym of the away team.

# Question 6

As requested, I'm going to pretend to make a hockey dataset. I will cut and paste the "data" from the assignment first:

```{r}
name <-
  c(
    "Wayne Gretzky",
    "Gordie Howe",
    "Jaromir Jagr",
    "Brett Hull",
    "Marcel Dionne",
    "Phil Esposito" ,
    "Mike Gartner",
    "Alex Ovechkin",
    "Mark Messier" ,
    "Steve Yzerman"
  )

goals <- c(894, 801, 766, 741, 731, 717, 708, 700, 694, 692)

year_started <- c(1979, 1946, 1990, 1986, 1971, 1963, 1979,
                  2005, 1979, 1983)
```

It looks like that actually worked. I have `goals`, `name`, and `year_started` in my environment now.

Now I will put these together in a tibble.

```{r}
hockey_data <- tibble(
  name = name,
  goals = goals,
  year_started = year_started
)

glimpse(hockey_data)
```

Looks like it worked!
