---
title: "Data Documentation"
author: "Nicolas Restrepo"
output: 
  html_document: 
    toc: true
    theme: united
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Here, I will keep track of some cool datasets to use in the course. 

## UN Votes 

This is a dataset that provides information about how countries have voted in the UN National Assembly. The data comes neatly bundled in a package that you can get by running `install.packages("unvotes")`.

The package contains three datasets: 

1) `un_votes` contains all the decisions for each resolution and each country. 

```{r}
# Load the packages
library(unvotes)
library(tidyverse)

# Eagle eye view of the data
glimpse(un_votes)
```

2) `un_roll_calls` includes information about each roll call: what the vote was about, when it happened, and what was being discussed: 

```{r}
glimpse(un_roll_calls)
```
3) `un_roll_call_issues` relates each vote to broader themes. 

```{r}
# View data 
glimpse(un_roll_call_issues)
# How many votes per issue? 
un_roll_call_issues %>% 
  count(issue, sort = T)
```

## Causact 

The `causact` package is designed to help with the construction and use of causal models. However, it also contains some interesting datasets. Let's highlight a couple. 

As usual, you can get this package by running `install.packages("causact")`

### Baseball

`baseballData` contains the final scores of 12,145 baseball games played in 2010. 

```{r}
library(causact)
glimpse(baseballData)
```
### Corruption 

If you are more into international relations and development, `corruptDF` contains information about the human development index and the corruption perception index for different countries. 

```{r}
glimpse(corruptDF)
```

### NTC Tickets

`ticketsDF` contains observations of the number of tickets written by NYC precincts each day between 2014 and 2015.

```{r}
glimpse(ticketsDF)
```

## Malaria 

If you are interested in public health, this could be of interest. In the folder, you will find some datasets compiled from the `malariaAtlas` package. There are three datasets here. The first contais the incidence of Malaria for each country each year. 

```{r}
df <- read_csv("malaria_incidence.csv")
glimpse(df)
```
The second dataset contains the deaths of malaria standardized across all ages for a given country-year. 

```{r}
df <- read_csv("malaria_deaths_standard.csv")
glimpse(df)
```
Finally, we have a dataset of malaria deaths but among specific age groups. 

```{r}
df <- read_csv("malaria_deaths_age_specific.csv")
glimpse(df)
```

## NFL Stats (2000-2017)

If you are interested in Football, we have a a dataset of player stats in the NFL from 2000 to 2017. 

```{r}
df <- read_csv("nfl_stats.csv")
glimpse(df)
```

Maybe you are interested in how the salaries for different positions have evolved across the years. We have you covered. 

```{r}
df <- read_csv("nfl_salaries.csv")
glimpse(df)
```

## California Fires 

Maybe you are interested in climate change. Then, the number of California fires across years, and the damaged they have caused, might be of interest. 

```{r}
df <- read_csv("fires.csv")
glimpse(df)
```

## TV & Movie Ratings 

We also have a dataset that contains TV & movie ratings taken from IMDB. The dataset covers titles released between 2000 and 2019. 

```{r}
df <- read_csv("tv_ratings.csv")
glimpse(df)
```

## Incarceration Trends

Here's a data about the people incarcerated at the county level from 1970 to 2015. This is a big dataset so it might take a while to load. 

```{r}
df <- read_csv("incarceration_trends.csv")
glimpse(df)
```

