---
title: "DV 4-5 Demo"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Make sure you have an up-to-date version of **toRvik**.

```{r, eval=FALSE}
devtools::install_github("andreweatherman/toRvik")
```

Load libraries.

```{r, message=FALSE}
library(tidyverse)
library(toRvik)
```

Load data.

```{r}
d <- bart_player_season(year = 2022,
                        stat = "box")
glimpse(d)
```

Let's deal with the weird height variable.

```{r}
d |> count(hgt)
```

Use separate to split the character `hgt` column into two new variables: `feet` and `inches`. Then we can make a new variable, `height`, which is the total height in inches.

```{r}
d <- d |> 
  separate(hgt,
           sep = "-",                    # the character to split on
           into = c("feet", "inches"),   # the two new variables
           convert = TRUE) |>            # make them numeric
  mutate(height = feet*12 + inches)      # create the new height variable
```

Here's the height distribution.

```{r}
p <- ggplot(d,
       aes(x = height,
           y = (..count../sum(..count..)) * 100 )) +    # manual y
  geom_histogram(binwidth = 1,
                 color = "white") +
  labs(y = "Percent of players",
       x = "Height in Inches") +
  scale_x_continuous(breaks = seq(60, 96, 6),
                     minor_breaks = NULL) +         # no minor gridlines
  scale_y_continuous(breaks = 0:11,
                     minor_breaks = NULL) +         # no minor gridlines
  theme_minimal()

p
```

We want to see it separately by position. But this doesn't work quite right.

```{r}
p + facet_wrap(~ pos)
```

As I have been saying, it's usually easier to do the data prep outside of the plot. Since height is discrete (an integer), I can use `geom_bar()` here instead.

```{r}
d |> 
  select(pos, height) |> 
  drop_na() |> 
  group_by(pos, height) |> 
  summarize(count = n()) |>   # this step will drop a group so it's just pos
  mutate(pct = (count / sum(count)) * 100) |>
  
  ggplot(aes(x = height ,
             y = pct )) +
  geom_bar(stat = "identity") +
  facet_wrap(~ pos) +
  scale_x_continuous(breaks = seq(60, 96, 6),
                     minor_breaks = NULL) +
  theme_minimal()
```

Let's put the scales on every facet.

```{r}

# data prep
d |> 
  select(pos, height) |> 
  drop_na() |> 
  group_by(pos, height) |> 
  summarize(count = n()) |>   # this step will drop a group so it's just pos now
  mutate(pct = (count / sum(count)) * 100) |>  # sending to plot

# plot
  ggplot(aes(x = height ,
             y = pct )) +
  geom_bar(stat = "identity") +
  facet_wrap(~ pos, scales = "free") +
  scale_y_continuous(limits = c(0, 30),
                     breaks = seq(0, 30, 5),
                     minor_breaks = NULL) +
  scale_x_continuous(limits = c(60, 90),
                     breaks = seq(60, 96, 6),
                     minor_breaks = NULL) +
  theme_minimal()
```

## Two dimensions

Let's use data from the TV ratings data you'll need for your homework.

```{r}
library(here)
tvd <- read_csv(here("Data", "tv_ratings.csv"))
```

```{r}
glimpse(tvd)
```

Let's look at the shows that have been around for at least 5 seasons

```{r}
tvd5 <- tvd |> 
  group_by(titleId) |> 
  mutate(max_seasons = max(seasonNumber)) |> 
  filter(max_seasons >= 5) |> 
  ungroup()

glimpse(tvd5)
```

```{r}
ggplot(tvd5,
       aes(x = seasonNumber,
           y = av_rating,
           group = titleId)) +
  geom_line(alpha = .2) +
  theme_minimal() +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none") 
```

```{r}
ggplot(tvd5,
       aes(x = seasonNumber,
           y = av_rating,
           group = titleId)) +
  geom_line(alpha = .2) +
  theme_minimal() +
  geom_smooth(aes(group = 1),
              se = FALSE) +
  theme(legend.position = "none") 
```

