---
title: "MD2 CB Demo"
author: "Stephen Vaisey"
date: "2022-09-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages

```{r}
library(tidyverse)
library(hoopR)
```

Load the data. I set `cache = TRUE` to keep from unnecessary redownloads.

```{r, cache=TRUE}
tb <- load_mbb_team_box()
```

Filter data so it's only Duke and only variables with `pct` in them. This will give me shooting accuracy data.

```{r}
duke <- tb |> 
  filter(team_short_display_name == "Duke") |> 
  select(game_date, contains("pct"))
glimpse(duke)
```

From the `glimpse()` above, it looks like the numbers are stored as character variables. You can see this with the `<chr>` indicator. I will proces the variables so they become numeric. This is tricky so just copy along here.

```{r}
duke <- duke |> 
  mutate(across(field_goal_pct:free_throw_pct,
                ~ as.numeric(.x)))
glimpse(duke)
```

OK looks good. Let's make some graphs.

```{r}
ggplot(duke,
       aes(x = free_throw_pct)) +
  geom_histogram()
```

The histogram produced by the code above is ugly. I am going to play with the settings to see if I can make something better.

```{r}
ggplot(duke,
       aes(x = free_throw_pct)) +
  geom_histogram(boundary = 40,
                 binwidth = 5,
                 color = "white") +
  labs(title = "Duke Free Throw % by Game, 2021",
       x = "Free Throw %",
       y = "Number of Games")
```


```{r}
ggplot(duke,
       aes(x = game_date,
           y = field_goal_pct)) +
  geom_line() +
  geom_point() +
  geom_smooth(se = FALSE)
```

