---
title: "MD3 Demo"
author: "Stephen Vaisey"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install new package (h/t Andrew Weatherman) for college basketball data. (Don't forget `eval = FALSE` in the code chunk!)

```{r, eval=FALSE}
install.packages("toRvik")
```

Let's load the packages we will need.

```{r}
library(toRvik)
library(tidyverse)
```

I'm going to grab player data by game in 2022. I'm going to add `cache = TRUE` to the chunk options so I don't have to re-download the data every time.

```{r, cache=TRUE}
pg <- bart_player_game(year = 2022,
                       stat = "box")
```

We are going to learn several `dplyr` verbs using this dataset. The main ones are:

-   `filter()`
-   `summarize()`
-   `group_by()`
-   `mutate()`
-   `arrange()`

Before we use these, let's `glimpse()` the data. (Note that `glimpse()` is also a verb!)

```{r}
glimpse(pg)
```

There are `r nrow(pg)` rows in this dataset. Each one corresponds to one player's performance in one game in the 2021-22 season.

There are lots of variables here we could use. For example:

-   `pts` is points scored
-   `fga` is field goals attempted
-   `fgm` is field goals made

These data come from all 32 Division I conferences. Let's use the `count()` verb to check this out.

```{r}
pg |> count(conf)
```

This might be too much. We can use the `filter()` verb to limit our exploration to the ACC.

```{r}
pg_acc <- pg |> 
  filter(conf == "ACC")
```

Now we have a new dataframe: `pg_acc`. Let's confirm that it only has teams from the ACC.

```{r}
pg_acc |> count(team)
```

We can also use `filter()` without creating a new dataframe. For example, we could filter the dataframe to just Duke and see how many games each player played. We use the `count()` variable to count the number of rows for each player.

```{r}
pg_acc |>                    # start with ACC player data
  filter(team == "Duke") |>  # keep Duke rows
  count(player)              # count player appearances  
```

Note that this is simply the number of rows in which each player appears in the dataset. Since each row is a game in which a player plays, the count is just the number of games a player played in.

This is great but it's probably not in the order we want. We can add the `arrange()` verb to the bottom of the pipeline to control this.

```{r}
pg_acc |>                    
  filter(team == "Duke") |>  
  count(player) |> 
  arrange(n)                 # sort by the number of appearances
```

If we want to sort it in _descending_ order, we wrap the ordering variable in `desc()`, like this:

```{r}
pg_acc |>                    
  filter(team == "Duke") |>  
  count(player) |> 
  arrange(desc(n))                 
```

Using `count()` is fine for a quick glance at frequencies. But using `summarize()` gives us more control at the cost of being a bit trickier to use at first. 

We can start by getting the exact same thing as `count()`. First we'll try something that **doesn't work as intended** and then we will look at something that _does_ work.

```{r}
pg_acc |> 
  filter(team == "Duke") |> 
  summarize(n = n())
```

What is happening here? This counts the total number of games played by _all_ Duke players in 2021-22. If we want it separately for each player, we need to use one of the most useful `dplyr` verbs: `group_by()`. This verb performs _every subsequent operation in the pipeline_ separately for each individual value of that variable. Here we're going to `group_by(player)` to calculate the number of appearances separately by player.

```{r}
pg_acc |> 
  filter(team == "Duke") |> 
  group_by(player) |> 
  summarize(n = n()) |> 
  arrange(desc(n))
```

This is more involved than `count()` but now we're not stuck with just one new variable. We can create as many new variables as we want. We have to make up a new variable name and use _functions_ that will create that new variable out of an existing variable. 

We could do this using a lot of different functions. Let's look at some examples.

```{r}
pg_acc |> 
  filter(team == "Duke") |> 
  group_by(player) |> 
  summarize(n = n(),
            mean_pts = mean(pts),   # average per game
            total_pts = sum(pts),   # total from all games
            max_pts = max(pts)) |>  # max pts in single game
  arrange(desc(mean_pts))
```

We used three different functions here: `mean()`, `sum()`, and `max()` and used them to create new variables. (We could also use `arrange()` if we wanted to sort the table by any of these but you've seen that already.)

Of course, we didn't have to group by player. Let's group by "experience" (class) for the whole ACC instead:

```{r}
pg_acc |> 
  group_by(exp) |> 
  summarize(total_pts = sum(pts)) |> 
  arrange(desc(total_pts))
```

This is nice and all, but why not plot it? We can pipe a tibble we've created directly into `ggplot()` if we want. This will often give us a little more control over what we want to see.

```{r}
pg_acc |> 
  group_by(exp) |> 
  summarize(total_pts = sum(pts)) |> 
  ggplot(aes(x = reorder(exp, total_pts),  # graph starts here
             y = total_pts)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(x = "Class",
       y = "Total points",
       title = "Points scored by class",
       subtitle = "ACC Men's Basketball, 2021-22 Season")
```

I saw a lot of you were fighting with `ggplot()` to reorder x-axes on, for example, bar graphs. This is normal; I often fight with them myself. One way to get around this is to make the tibble you want first before sending it to `ggplot()`.

Here's an example. We can use `mutate()` to create the variables we want without having to try to get them right within the `ggplot()` syntax. In this example, we are using `case_when()` to map an old variable (`exp`) to a new variable (`class`). We are also creating a numeric version (`class_num`) to make it easier to sort the x-axis the way we want. Be warned: `case_when()` is tricky but really powerful once you get it figured out.

```{r}
pg_acc2 <- pg_acc |>
  mutate(
    class = case_when(
      exp == "Fr" ~ "1st Year",
      exp == "So" ~ "Sophomore",
      exp == "Jr" ~ "Junior",
      exp == "Sr" ~ "Senior"),
    class_num = case_when(
      exp == "Fr" ~ 1,
      exp == "So" ~ 2,
      exp == "Jr" ~ 3,
      exp == "Sr" ~ 4))

glimpse(pg_acc2)
```

We now have two new columns visible at the bottom of this output: `class` and `class_num`. This gives us new, easier to plot with, versions of the `exp` variable. We can do even more data transformations to get our data ready to send to `ggplot()`:

```{r}
graph_data <- pg_acc2 |> 
  group_by(class) |> 
  summarize(class_num = min(class_num),  # constant within groups!
            total_pts = sum(pts))
```

This prep makes the subsequent `ggplot()` easier to work with because we have everything exactly where we want it without having to do it inside the `ggplot()` function.

```{r}
ggplot(graph_data,
       aes(x = reorder(class, class_num),  # reorder by num version
           y = total_pts)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(x = "Class",
       y = "Total points",
       title = "Points scored by class",
       subtitle = "ACC Men's Basketball, 2021-22 Season")
```

OK let's take a step back. Not all uses of `mutate()` have to be as complicated as `case_when()`. Consider the following:

```{r}
pg_acc |> 
  select(player, date, fga, fgm) |> 
  head(10)
```

This is the first 10 games the first player played in the ACC data. We see how many shots this player attempted and how many he made in each game. We could turn this into a percentage easily using `mutate()`.

```{r}
pg_acc |> 
  select(player, date, fga, fgm) |> 
  mutate(fg_pct = 100*(fgm/fga)) |> 
  head(10)
```

Maybe this is too bouncy game by game. We could instead calculate the field goal percentage for each player's season using `group_by()` and `summarize()`.

```{r}
pg_acc |> 
  select(player, date, fga, fgm) |> 
  group_by(player) |> 
  summarize(tot_fga = sum(fga),
            tot_fgm = sum(fgm)) |> 
  mutate(fg_pct = 100*(tot_fgm/tot_fga)) |> 
  head(10)
```

We have two problems here. First, we have some players who didn't attempt any shots so they get `NaN` for their shooting percentage. Second, we probably don't want to see this in alphabetical order. So we can use `arrange(desc())` to sort it from highest to lowest.

```{r}
pg_acc |> 
  select(player, date, fga, fgm) |> 
  group_by(player) |> 
  summarize(tot_fga = sum(fga),
            tot_fgm = sum(fgm)) |> 
  mutate(fg_pct = 100*(tot_fgm/tot_fga)) |> 
  arrange(desc(fg_pct)) |> 
  head(10)
```

But wait. Something still seems off here. Are these the best shooters from last season? Why not?

Maybe we should `filter()` the data so we only see the players who've taken at least 100 shots. What would that look like?

```{r}
pg_acc |> 
  select(player, date, fga, fgm) |> 
  group_by(player) |> 
  summarize(tot_fga = sum(fga),
            tot_fgm = sum(fgm)) |> 
  mutate(fg_pct = 100*(tot_fgm/tot_fga)) |>
  filter(tot_fga > 100) |> 
  arrange(desc(fg_pct)) |>
  head(10)
```

I realize we're mixing up 2-point and 3-point shooting here but you get the idea.